The Heavy Photon Search Test Detector

### 2014-04-08 hps-testrun-nim.tex
Edits from Tim

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the Test Run apparatus 
and readout electronics and its performance. In this setting, a heavy photon can be identified as a 
narrow\ee  mass peak above the trident background or as a decay vertex displaced from
the production target, so charged particle tracking and vertexing are needed for its detection. In the 
HPS Test run, charged particles are measured
with a compact forward silicon microstrip vertex/tracker inside a dipole magnet. Electromagnetic 
showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test run utilizes  high-rate readout and data acquisition electronics and a fast trigger to
perfectly exploit the essentially 100\% duty cycle of the CEBAF accelerator.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a "hidden sector" or "dark" photon, is a particle with mass 
$10-1000$~MeV which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous Beyond Standard Model theories `\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical anomalies.  Heavy photons  couple directly to  hidden sector 
particles with "dark" or "hidden sector" charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. [REFERENCE]. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime's become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

@@

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test apparatus was designed to run in Hall~B at JLab using the CEBAF electron beam, a 
499~MHz beam, at energies of 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\textrm{A}^\prime}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a "wall of flame" which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a "dead zone", the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  PbWO$_{4}$ crystal calorimeter located just downstream 
of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < XXXX$ & 12800 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < XXXX$ & 442 & 12 & 4~ns & 1 & $1.3\times1.3$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

@@

\section{The HPS Test Run Beamline}
The HPS Test run studied multiple Coulomb scattering of electrons and positrons from 
bremsstrahlung photons produced in the Hall B tagged photon facility. 
Figure~\ref{fig:hpstest_layout} shows the layout of 
the setup on the beam line. The SVT was installed inside the Hall B pair 
spectrometer magnet vacuum chamber with the ECal mounted downstream of it. Both the 
SVT and the ECal were retracted off the beam plane compared to nominal electron beam runing to 
allow clean passage of the photon beam through the system. 
\begin{figure}[]
    \includegraphics[width=0.5\textwidth]{figures/HPS_dimensions}
\caption{\small{Layout of the HPS parasitic run.} }
\label{fig:hpstest_layout}
\end{figure}

The photon beam was generated in the interaction of $5.5$~GeV electrons with a $10^{-4}~X_0$ 
gold radiator located $\approx~9$~m upstream of the PS. The primary beam and scattered 
electrons are deflected away from detectors by the dipole magnet of the photon tagging system. 
During the dedicated HPS Test run period, the collimated (6.4~mm diameter), photon beam passes 
through the PS pair converter gold foil and later the HPS system as illustrated in 
Fig.~\ref{fig:schematic_testrun_vs_erun}.
\begin{figure}[]
\includegraphics[ width=7cm]{figures/photon_vs_electron_beam_schematic.png}
\caption{\small{Illustrative comparison of HPS Test run photon beam compared to the HPS electron 
beam.}}
\label{fig:schematic_testrun_vs_erun}
\end{figure}
The PS pair converter was located $\approx~77$~cm upstream of the first layer of the SVT.

 Data was taken on three different converter thicknesses with photon fluxes between 
 1.1-2.6$\times10^{8}$/s ($0.55<$E$_\gamma<5.5$~GeV) at beam currents varying between 
 30-90~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the test run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime's is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10 cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5 mm from the center of the beam, 
defining the 15 mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$ electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

\subsection{Layout}
The layout of the SVT is summarized in Tab.~\ref{tab:trk} and rendered in Fig.~\ref{fig:tracker_model}. 
Each of the layers is comprised of a pair of closely-spaced silicon microstrip sensors mounted back-
to-back to form a module. A 100~mrad stereo angle is used in the first three layers to provide higher-
resolution 3D space points for vertexing.  Using 50~mrad in the last two layers breaks the tracking 
degeneracy of having five identical layers and minimizes fakes from ghost hits to improve pattern 
recognition. Altogether, the SVTcomprises 20 sensors for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\footnotesize
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{\small Layout of the SVT.}
}
\label{tab:trk}
\vspace*{-15mm}
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{\small A rendering of the SVT showing the modules on their support plates held by the 
hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the 
hybrid readout boards in green. The beam enters from the right through a vacuum box with flanges 
for services. }
\label{fig:tracker_model}
%\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The SVT is built in two separate halves that are mirror reflections of one another about the plane of 
the nominal electron beam.  Each half consists of five modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. 
The two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond 
the last layer that defines the nominal spacing between the upper and lower halves of the tracker.  A 
shaft attached to each support plate in front of layer 1 extends upstream and connects to a linear shift 
that transfers motion into the vacuum box through bellows to open and close the two halves around 
the dead zone. The C-support is mounted to an aluminum baseplate that defines the position of the 
SVT with respect to the vacuum chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both 
completed detector halves prior to final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
%\vspace*{-5mm}
\end{figure}
%=======================

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 can capture two successive samples of three 
of the output of the shaper at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx25$~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) is used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards, outside the tracking volume, with a 
short twisted-pair pigtail cable to provide power and configuration and signal readout. Along with a 
single sensor, these are glued to a polyamide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyamide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the backside of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

\subsection{Alignment}
The SVT was aligned using a combination of optical, laser and touch probe surveys at SLAC and JLab. 
The optical survey of individual modules with precision of a few $\mu$m are combined with a 
touch-probe survey of the overall SVT support structure, with 25-100~$\mu$m precision, to locate the 
silicon sensor layers with respect to the support plates and the mechanical survey balls on the base 
plate. After full assembly and installation of the SVT at JLab, a mechanical survey of the SVT base plate 
position inside the pair spectrometer vacuum chamber is used to determine the global position of the 
SVT with respect to CEBAF beam line. The resulting survey-based alignment has the position of the 
silicon sensors correct to within a few hundred microns measured from tracks in the Test run data. 
A more sophisticated global track-based alignment technique to reach final alignment precision 
well below 50~$\mu$m is being developed.

@@

The electromagnetic calorimeter (ECal), installed downstream of the PS dipole magnet, performs two 
essential 
functions for the experiment: it provides a trigger signal to select what events to read out from the 
detector sub-systems and is used in the analysis to identify electrons and positrons. 
The technology and design choices are largely driven by the need for a compact forward design 
covering the SVT \Aprime acceptance and able to fully absorb 
electrons and positrons with energy between $0.5$-$6.5$~GeV, fine granularity and signal 
readout speed to handle 1~MHz/cm$^{2}$ of electromagnetic background and remain operable 
after a radiation dose larger than $X$~MRad. The lead-tungstate (PbOW$_{4}$) crystal inner 
calorimeter of the CLAS detector~\cite{2008erhm.conf..421N} in operation since 2005 in Hall B meets 
all the requirements set by HPS. The modules from this calorimeter have been subsequently 
repurposed for HPS.

@@

During the Test run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure}[]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
%\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view of the ECal looking downstream.}
\label{fig:ecal}
\end{center}
\end{figure}
}

>>>>

\begin{abstract}
The  Heavy Photon Search (HPS), an experiment to search for a hidden sector photon in fixed target 
electroproduction, is preparing for installation at the Thomas Jefferson National Accelerator Facility 
(JLab) in the Fall of 2014.  As the first stage of this project, the HPS Test Run apparatus was 
constructed and operated in 2012 to demonstrate the experiment's technical feasibility and to  confirm 
that the trigger rates and occupancies are as expected. This paper describes the Test Run apparatus 
and readout electronics and its performance. In this setting, a heavy photon can be identified as a 
narrow\ee  mass peak above the trident background or as a decay vertex displaced from
the production target, so charged particle tracking and vertexing are needed for its detection. In the 
HPS Test Run, charged particles are measured
with a compact forward silicon microstrip vertex/tracker inside a dipole magnet. Electromagnetic 
showers are detected in a PbW0$_{4}$ 
crystal calorimeter situated behind the magnet used to trigger the experiment and identify 
electrons and positrons. Both detectors are placed close to the 
beam line and split top-bottom. This arrangement  provides sensitivity to low-mass heavy photons,  
allows clear  passage of the unscattered beam, 
and avoids the spray of degraded electrons coming from the target. 
The discrimination between prompt and displaced \ee pairs requires the first layer of silicon 
sensors be placed only 10~cm downstream of the target. The expected signal is small, and the 
trident background huge, so the experiment requires very large statistics. Accordingly, the
HPS Test Run utilizes  high-rate readout and data acquisition electronics and a fast trigger to
perfectly exploit the essentially 100\% duty cycle of the CEBAF accelerator.
\end{abstract}

@@

%% main text
\section{Introduction}
\label{introduction}
The heavy photon (\Aprime{}), aka a "hidden sector" or "dark" photon, is a particle with mass 
$10-1000$~MeV which couples weakly to electric charge by mixing with the Standard Model 
photon~\cite{Holdom:1985ag}. Consequently, it can be radiated by electrons and subsequently decay 
into \ee pairs, albeit at rates far below those of QED trident processes. Heavy photons have been 
suggested by numerous Beyond Standard Model theories `\cite{Essig:2013lka}, to explain the 
discrepancy 
between theory and experiment of the muon's $g-2$ ~\cite{Pospelov:2008zw}, and as a possible 
explanation of recent astrophysical anomalies.  Heavy photons  couple directly to  hidden sector 
particles with "dark" or "hidden sector" charge; these particles
could constitute all or some of the dark matter~\cite{ArkaniHamed:2008qn}. 
Current phenomenology highlights the $20-1000$~MeV/c$^{2}$ mass range, and 
suggests that  the coupling to electric charge, $\epsilon$e, has $\epsilon$ in the range of 
$10^{-3} -10^{-5}$. [REFERENCE]. This range of parameters makes \Aprime{} searches viable in 
medium energy fixed target electroproduction~\cite{Bjorken:2009mm}, but requires large data sets and 
good mass resolution to identify a small mass peak above the copious QED background. At small 
couplings, \Aprime{} become long-lived, so detection of a displaced decay vertex can reject the prompt 
QED background and boost experimental sensitivity.

@@

\section{Detector Overview}
\label{detector}
{\small
\begin{figure*}[t]
\begin{center}
    \includegraphics[width=0.8\textwidth]{figures/hps_testrun_rendering}
\caption{Rendering of the HPS Test apparatus installed on the beam line.}
\label{fig:testrundetector}
\end{center}
\end{figure*}
}
The HPS Test apparatus was designed to run in Hall~B at JLab using the CEBAF electron beam, a 
499~MHz beam, at energies of 2.2 and 6.6~GeV and currents between 200 and 600~nA.  
The overall design of the experiment follows from the kinematics of \Aprime{} production which 
typically results in a final state particle within a few degrees of the incoming beam, especially at low 
$m_{\Aprime}$. Detectors must therefore be placed close to the beam. 
The intense electron beam enlarges downstream after multiple scattering in the target and electrons 
which have radiated in the target disperse horizontally in the field of the analyzing magnet. Together 
they constitute a "wall of flame" which must be completely avoided. Accordingly, 
the apparatus is split vertically to avoid a "dead zone", the region within $\pm 15$~mrad of the beam 
plane. In addition, the beam is transported in vacuum through the tracker to minimize beam-gas 
interaction backgrounds. Even with these precautions, the occupancies of sensors near the beam 
plane are high, dominated by the
multiple Coulomb scattering of the primary beam, so high rate detectors, a fast trigger, 
and excellent time tagging are required to minimize their impact.  
The trigger comes from a highly-segmented  PbWO$_{4}$ crystal calorimeter located just downstream 
of the dipole magnet. 
 %\begin{multicols}{2}
\begin{center}
\begin{table*}[t]
{\small
\caption{Overview of the coverage, segmentation and performance of the HPS Test detector. }
\begin{tabular}{lccccccc}
\hline 
System & Coverage & \# channels & ADC & Time resolution & \# layers & Segmentation & Performance \\
 & (mrad) &  & (bit) & (ns) &  &  &  \\
\hline
SVT & $15<\theta_{y} < XXXX$ & 12800 & 14 & $\approx 2$~ns & 5  & $\approx 120~\mu$m $r-\phi$ & $\sigma_{d0,y}  \approx 100~\mu$m \\
& &  &  &  & (stereo layers) & $\approx 6~\mu$m $z$ & $\sigma_{d0,x} \approx 300~\mu$m \\
& &  &  &  &  &  & $\sigma_{d0,z}\approx 1$~mm \\
\hline
ECal & $15<\theta_{y} < XXXX$ & 442 & 12 & 4~ns & 1 & $1.3\times1.3$~cm$^2$  & $\sigma(E)/E \approx 4.5\%$ \\ 
 &  &  &  &  &  & $1.6\times1.6$~cm$^2$  &  \\ 
\hline
\end{tabular}
\label{tab:detector-overview}
}
\end{table*}
\end{center}
%\end{multicols}
A rendering of the apparatus installed on the beam line is shown in 
Fig.~\ref{fig:testrundetector} and an overview of the coverage, segmentation and performance is 
given in Tab.~\ref{tab:detector-overview}.

@@

\section{The HPS Test Run Beamline}
The HPS Test Run studied multiple Coulomb scattering of electrons and positrons from 
bremsstrahlung photons produced in the Hall B tagged photon facility. 
Figure~\ref{fig:hpstest_layout} shows the layout of 
the setup on the beam line. The SVT was installed inside the Hall B pair 
spectrometer magnet vacuum chamber with the ECal mounted downstream of it. Both the 
SVT and the ECal were retracted off the beam plane compared to nominal electron beam runing to 
allow clean passage of the photon beam through the system. 
\begin{figure}[]
    \includegraphics[width=0.5\textwidth]{figures/HPS_dimensions}
\caption{\small{Layout of the HPS parasitic run.} }
\label{fig:hpstest_layout}
\end{figure}

The photon beam was generated in the interaction of $5.5$~GeV electrons with a $10^{-4}~X_0$ 
gold radiator located $\approx~9$~m upstream of the PS. The primary beam and scattered 
electrons are deflected away from detectors by the dipole magnet of the photon tagging system. 
During the dedicated HPS Test Run period, the collimated (6.4~mm diameter), photon beam passes 
through the PS pair converter gold foil and later the HPS system as illustrated in 
Fig.~\ref{fig:schematic_testrun_vs_erun}.
\begin{figure}[]
\includegraphics[ width=7cm]{figures/photon_vs_electron_beam_schematic.png}
\caption{\small{Illustrative comparison of HPS Test Run photon beam compared to the HPS electron 
beam.}}
\label{fig:schematic_testrun_vs_erun}
\end{figure}
The PS pair converter was located $\approx~77$~cm upstream of the first layer of the SVT.

 Data was taken on three different converter thicknesses with photon fluxes between 
 1.1-2.6$\times10^{8}$/s ($0.55<$E$_\gamma<5.5$~GeV) at beam currents varying between 
 30-90~nA and repeated with the reverse field setting of the PS dipole magnet.
\begin{table}[]
\begin{center}
{\small
\begin{tabular}{|c|c|c|}
\hline
Converter thickn. & Duration &  $e^-$ on radiator \\
 (\%$X_0$) & (s) & ($\mu$C)    \\   
\hline
%\hline
1.6   & 911 &   24.4 \\ %24385.9     \\ %27 nA
%\hline
0.18   & 2640 &   193.5 \\ % 193508.9  \\  %73 nA
%\hline
0.45  & 2149 &     140.7 \\ %  140709.9  \\ %65.5 nA
%\hline
0    & 1279  &   88.1 \\ %88079.6  \\
\hline
\end{tabular}
}
\caption{Measured integrated currents for the dedicated photon runs.}
\label{tab:currents}
\end{center}
\end{table}
The photon beam line during the Test Run produced a relatively large fraction of pairs 
originating upstream of the converter position. This contribution was measured during data taking 
with ``empty'' converter runs i.e. removing the converter but with all other conditions the same. 
The runs taken during the time dedicated to HPS Test is summarized in Tab.~\ref{tab:currents}.

@@

The design of the SVT is primarily driven by direct physics requirements and constraints from the 
environment at the interaction region. The \Aprime{} decay products 
have momenta in the range of 1~GeV, so multiple scattering dominates mass and vertexing 
uncertainties for any possible material budget, so the SVT must minimize the amount of 
material in the tracking volume. The signal yield for long-lived \Aprime{} is very small, so 
the rejection of prompt vertexes must be exceedingly pure, on the order of $10^{-7}$, in order to 
eliminate all prompt backgrounds. To achieve the required vertexing performance the first layer of the 
SVT must be placed no more than about 10 cm downstream of the target. At that distance, it is found 
that the active region of a sensor can be placed as close to 1.5 mm from the center of the beam, 
defining the 15 mrad ``dead zone" mentioned previously, to maximize low-mass \Aprime{} acceptance 
with decay products nearly collinear with the beam axis. At the edge of this ``dead zone," the 
radiation dose approaches $10^{15}$ electrons/cm$^2$/month, or roughly 
$3 \times 10^{14}$~\fluenceunit{}/month~\cite{Rashevskaya:2002nd}, 
requiring the sensors to be actively cooled.
% and~\ref{fig:rates}.
%===================
%\begin{figure}[]
%\includegraphics[width=7cm]{figures/radiation}
%\label{fig:radiation}
%\caption{{\small Simulation of the electron fluence 10~cm downstream of the target in the active region 
%of the SVT ($|Y|>0.15~$mm) looking downstream; electrons bend towards $+x$.}}
%\end{figure}
Meanwhile, very low-energy delta rays from 
beam-gas interactions multiply the density of background hits, so the SVT must operate inside the 
beam vacuum.  Finally, in order to protect the sensors, the detector must be movable so that it can be 
retracted during periods of uncertain beam conditions.

\subsection{Layout}
The layout of the SVT is summarized in Tab.~\ref{tab:trk} and rendered in Fig.~\ref{fig:tracker_model}. 
Each of the layers is comprised of a pair of closely-spaced silicon microstrip sensors mounted back-
to-back to form a module. A 100~mrad stereo angle is used in the first three layers to provide higher-
resolution 3D space points for vertexing.  Using 50~mrad in the last two layers breaks the tracking 
degeneracy of having five identical layers and minimizes fakes from ghost hits to improve pattern 
recognition. Altogether, the SVT has 20 sensors for a total of 12780 readout channels. 
%=======================
\begin{center}
\begin{table}[ht]
{\footnotesize
\begin{tabular}{lccccc}   
\hline \hline 
    Layer & 1 & 2 & 3 & 4 & 5 \\      
\hline
    $z$ from target (cm)  & 10 & 20 & 30 & 50 & 70  \\ 
    Stereo angle (mrad)  & 100 & 100 & 100 & 50 & 50 \\ 
    Bend res. ($\mu$m)  & $\approx$60 & $\approx$60 & $\approx$60 & $\approx$120 & $\approx$120  \\ 
    Non-bend res. ($\mu$m)  & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6 & $\approx$6  \\ 
    \# of sensors  & 4 & 4 & 4 & 4 & 4  \\ 
    Dead zone (mm) & $\pm1.5$  & $\pm3.0$  & $\pm4.5$  & $\pm7.5$  & $\pm10.5$  \\ 
    Power cons. (W) & 6.9 & 6.9 & 6.9 & 6.9 & 6.9 \\
\hline \hline
\end{tabular}
\caption{\small Layout of the SVT.}
}
\label{tab:trk}
\vspace*{-15mm}
\end{table}
\end{center}
%=======================
\begin{center}
\begin{figure}[htp]
\includegraphics[width=7cm]{figures/HPS_nochamber}
\caption{\small A rendering of the SVT showing the modules on their support plates held by the 
hinged C-support on the left and the motion levers on the right. The sensors are shown in red and the 
hybrid readout boards in green. The beam enters from the right through a vacuum box with flanges 
for services. }
\label{fig:tracker_model}
%\vspace*{-5mm}
\end{figure}
\end{center}
%=======================
The SVT is built in two separate halves that are mirror reflections of one another about the plane of 
the nominal electron beam.  Each half consists of five modules mounted on a support plate that 
provides services to the modules and allows them to be moved as a group relative to the dead zone. 
The two halves of the tracker are connected to hinges mounted on a C-shaped support just beyond 
the last layer that defines the nominal spacing between the upper and lower halves of the tracker.  A 
shaft attached to each support plate in front of layer 1 extends upstream and connects to a linear shift 
that transfers motion into the vacuum box through bellows to open and close the two halves around 
the dead zone. The C-support is mounted to an aluminum baseplate that defines the position of the 
SVT with respect to the vacuum chamber. Figure~\ref{fig:tracker_halves} shows a photograph of both 
completed detector halves prior to final assembly. 
%=======================
\begin{figure}[htp]
\begin{center}
    \includegraphics[width=7cm]{figures/2012-101-PHOTON-DETECTOR-001}
\caption{\small{Both halves of the HPS Test SVT after final assembly at SLAC.  The cooling manifolds and 
integrated cable runs are clearly seen.} }
\label{fig:tracker_halves}
\end{center}
%\vspace*{-5mm}
\end{figure}
%=======================

@@

%\subsubsection{Front-end Electronics}
Despite having only small spots with very high occupancy (up to 4~MHz/mm$^2$) closest to the primary 
beam, the rates are still high and lowering the peak occupancy to 
approximately 1\% for tracking requires a trigger window and hit time tagging of roughly 8~ns. The 
ECal readout and trigger described in Sec.~\ref{sec:fadc} can achieve such resolution. To reach this 
performance the sensors for the SVT are readout by the APV25 ASIC developed for the CMS 
experiment at CERN~\cite{French:2001xb}. The APV25 can capture successive samples of the shaper output 
in groups of three at a sampling rate of approximately 40~MHz.  By fitting the known 
$CR$-$RC$ shaping curve to these samples, the initial time of the hit can be determined to a precision 
of 2~ns for S/N$\approx25$~\cite{Friedl:2009zz}.  For electron beam running, six-sample readout and 
the shortest possible shaping time (35~ns) is used to best distinguish hits that overlap in time.
The APV25 ASICs are hosted on simple FR4 hybrid readout boards, outside the tracking volume, with a 
short twisted-pair pigtail cable to provide power and configuration and signal readout. Along with a 
single sensor, these are glued to a polyamide-laminated carbon fiber composite backing making 
up a half-module. A window is machined in the carbon fiber leaving only a frame around the periphery 
of the silicon to minimize material. A 50~$\mu$m sheet of polyamide is laminated to the surface of the 
carbon fiber with 1~mm overhang at all openings to ensure good isolation between the backside of the 
sensor, carrying high-voltage bias, and the carbon fiber which is held near ground.

@@

\subsection{Alignment}
The SVT was aligned using a combination of optical, laser and touch probe surveys at SLAC and JLab. 
The optical survey of individual modules with precision of a few $\mu$m are combined with a 
touch-probe survey of the overall SVT support structure, with 25-100~$\mu$m precision, to locate the 
silicon sensor layers with respect to the support plates and the mechanical survey balls on the base 
plate. After full assembly and installation of the SVT at JLab, a mechanical survey of the SVT base plate 
position inside the pair spectrometer vacuum chamber is used to determine the global position of the 
SVT with respect to CEBAF beam line. The resulting survey-based alignment has the position of the 
silicon sensors correct to within a few hundred microns measured from tracks in the Test Run data. 
A more sophisticated global track-based alignment technique to reach final alignment precision 
well below 50~$\mu$m is being developed.

@@

The electromagnetic calorimeter (ECal), installed downstream of the PS dipole magnet, performs two 
essential 
functions for the experiment: it provides a trigger signal to select what events to read out from the 
detector sub-systems and is used in the analysis to identify electrons and positrons. 
The technology and design choices are largely driven by the need for a compact forward design 
covering the SVT \Aprime{} acceptance and able to fully absorb 
electrons and positrons with energy between $0.5$-$6.5$~GeV, fine granularity and signal 
readout speed to handle 1~MHz/cm$^{2}$ of electromagnetic background and remain operable 
after a radiation dose larger than $X$~MRad. The lead-tungstate (PbOW$_{4}$) crystal inner 
calorimeter of the CLAS detector~\cite{2008erhm.conf..421N} in operation since 2005 in Hall B meets 
all the requirements set by HPS. The modules from this calorimeter have been subsequently 
repurposed for HPS.

@@

During the Test Run, both halves were held in place by four vertical bars attached to an above rail, 
placing the front face of the crystals 147~cm from the upstream edge of the magnet and with a 
8.7~cm gap between the innermost edge of the crystals in the two halves.
{\small
\begin{figure}[]
\begin{center}
\includegraphics[width=0.45\textwidth]{figures/ECal}
%\includegraphics[width=0.45\textwidth]{figures/ecal_picture_beamline}
\caption{Rendered layout view of the ECal looking downstream.}
\label{fig:ecal}
\end{center}
\end{figure}
}
